---
title: "Bayesian Classifier"
author: "Jaime Díaz-Trechuelo Sánchez-Moliní"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
#Some of this packages might need to be installed first
library(ggplot2)
library(latex2exp)
library(ggrepel)
library(matlib)
```

## The Bayes' Classifier Regions
This part of the code gives the `BayesClassifier` function which asks for the means and variances of two bivariate normal distributions. It then represents the 2-dimensional critical regions associated with these random variables.
Mathematically we have:
Let $X$ and $Y$ be two random variables, 2 and 1 dimensional respectively s.t.
$$
(X|Y=k) \sim \mathcal{N}_k(\mu_k,\Sigma_k)  \quad k=1,2
$$
Our objective is to give a rule so that given an observation of $X$, noted $X_i$ we'll infere wether $X_i$ is associated to $Y=1$ or $Y=2$. One way to do this is with the Bayes' Classifier.

Let's now go over the parameters of this function:

`mu11` is the value of $(\mu_1)_1$ 

`mu12` is the value of $(\mu_1)_2$ 

`mu21` is the value of $(\mu_2)_1$

`mu22` is the value of $(\mu_2)_2$

`S11` is the value of $(\Sigma_1)_{11}$

`S12` is the value of $(\Sigma_1)_{22}$

`S1D` is the value of $(\Sigma_1)_{12} = (\Sigma_1)_{21}$

`S21` is the value of $(\Sigma_2)_{11}$

`S22` is the value of $(\Sigma_2)_{2}$

`S2D` is the value of $(\Sigma_2)_{12} = (\Sigma_2)_{21}$

If any of this values isn't specified then function assumes the value is taken as $0$
`margins` is a 4 element vector with the limits of the graphic window to be created in the order $(x_{1,min},x_{1,max},x_{2,min},x_{2,max})$. By default, if omitted, `c(-5,5,-5,5)` is taken.
`labels`is a boolean variable which is set to `TRUE` by default, it determines wether or not the regions are labelled $R_1,R_2$.


```{r BayesianClassifier, warning=FALSE}

BayesClassifier <- 
function(mu11=0,mu12=0,mu21=0, mu22=0, S11=0,S12=0,S1D=0,S21=0,S22=0,S2D=0, margins = c(-5,5,-5,5), labels = TRUE) {
  mu1 <- c(mu11, mu12)
  mu2 <- c(mu21, mu22)
  Sigma1 <- matrix(c(S11, S1D, S1D, S12), nrow = 2, byrow = TRUE)
  Sigma2 <- matrix(c(S21, S2D, S2D, S22), nrow = 2, byrow = TRUE)
  
  if (all(Sigma1==Sigma2)) {
    b <- inv(Sigma1) %*% (mu1 - mu2)
    region_condition <- function(x, y) {
      v <-matrix(c(x, y), ncol = 1)
      return(t(v) %*% b > 0.5 * t((mu1 + mu2)) %*% b)
    }} else {
      B <- inv(Sigma1) - inv(Sigma2)
      b <- inv(Sigma1) %*% mu1 - inv(Sigma2) %*% mu2
      k <- 0.5 * (log(det(Sigma2) / det(Sigma1)) + t(mu2) %*% inv(Sigma2) %*% mu2 - t(mu1) %*% inv(Sigma1) %*% mu1)
      region_condition <- function(x, y) {
        v <- matrix(c(x, y), ncol = 1)
        return(-0.5 * t(v) %*% B %*% v + t(v) %*% b > -k)
      }
    }
  
  x <- seq(margins[1], margins[2], length.out = 350)
  y <- seq(margins[3], margins[4], length.out = 350)
  data <- expand.grid(x = x, y = y)
  in_region = c()
  for (i in 1:length(x)) {
    for (j in (1:length(y))){
      in_region[350*(i-1)+j] <- region_condition(x[i],y[j])
    }}
  data$in_region <- in_region
  
  
  plot <- ggplot(data, aes(x, y, fill = in_region)) +
    geom_tile() +
    scale_fill_manual(values = c(rgb(1, 0, 0, alpha = 0.4), rgb(0, 0, 1, alpha = 0.4)), na.value = "white") +
    ggtitle("Critical Regions") +
    theme(legend.position = "none") +
    ylab(TeX("$X_2$")) + xlab(TeX("$X_1$"))
  
  if (labels) {
    x0 <- 0
    y0 <- 0
    if (region_condition(x0, y0)) {
      plot <- plot + annotate("text", x = 0, y = 0, label = TeX("$R_1$"), parse = TRUE, col = "blue", size = 12)
      stop = 0
      for (i in 1:length(in_region)) {
        if (!(region_condition((data$x)[i],(data$y)[i]))){
          stop <- stop+1
          if (stop == 7175){
            plot <- plot +  annotate("text", x = data$x[i] , y = data$y[i], label = TeX("$R_2$"), parse = TRUE, col = "red", size = 12)
            break}}}
    } else {
      plot <- plot +
        annotate("text", x = 0, y = 0, label = TeX("$R_2$"), parse = TRUE, col = "red", size = 12)
      stop = 0
      for (i in 1:length(in_region)) {
        if ((region_condition((data$x)[i],(data$y)[i]))){
          stop <- stop+1
          if (stop == 7175){
            plot <- plot +  annotate("text", x = data$x[i] , y = data$y[i], label = TeX("$R_1$"), parse = TRUE, col = "blue", size = 12)
            break}}}
    }    
  }
  plot}

```

## Bayesian Classifier for data sets

In this case we are given a data set of $n$ observations of a variable that follows what we laid out before. The function `BayesClassifierForSets` returns a plot were the points or the dataset are colored based on where they'd be classified. 

All parameters are the same as in the previous function, with the same defaults. The only changes are that there's no `margins` parameter and there is a parameter `observed_data` which is the data set which we want to classify, there's no default value for this parameter.

```{r BayesianClassifierForSets}
BayesClassifierForSets <- function(collected_data,mu11=0,mu12=0,mu21=0, mu22=0, S11=0,S12=0,S1D=0,S21=0,S22=0,S2D=0,labels=TRUE) {
  (mu1 <- c(mu11, mu12))
  mu2 <- c(mu21, mu22)
  Sigma1 <- matrix(c(S11, S1D, S1D, S12), nrow = 2, byrow = TRUE)
  Sigma2 <- matrix(c(S21, S2D, S2D, S22), nrow = 2, byrow = TRUE)
  if (all(Sigma1==Sigma2)) {
    b <- inv(Sigma1) %*% (mu1 - mu2)
    region_condition <- function(x, y) {
      v <-matrix(c(x, y), ncol = 1)
      return(t(v) %*% b > 0.5 * t((mu1 + mu2)) %*% b)
    }} else {
      B <- inv(Sigma1) - inv(Sigma2)
      b <- inv(Sigma1) %*% mu1 - inv(Sigma2) %*% mu2
      k <- 0.5 * (log(det(Sigma2) / det(Sigma1)) + t(mu2) %*% solve(Sigma2) %*% mu2 - t(mu1) %*% solve(Sigma1) %*% mu1)
      region_condition <- function(x, y) {
        v <- matrix(c(x, y), ncol = 1)
        return(-0.5 * t(v) %*% B %*% v + t(v) %*% b > -k)
      }
    }
  in_region = c()
  for (i in 1:length(collected_data[,1])) {
    in_region[i] <- region_condition(collected_data[i,1],collected_data[i,2])
  }
  collected_data<- cbind(collected_data,(abs(in_region-2)))
  indices_1 <-collected_data[ , 3] == 1
  indices_2 <- collected_data[ , 3] == 2
  xlim <- c(min(collected_data[,1])-1,max(collected_data[,1])+1)
  ylim <- c(min(collected_data[,2])-1,max(collected_data[,2])+1)
  plot(collected_data[indices_1,1],collected_data[indices_1,2], col = 'blue',ylab = TeX("$X_2"), xlab = TeX("$X_1"), xlim = xlim, ylim = ylim)
  points(collected_data[indices_2,1],collected_data[indices_2,2], col = 'red')
  title("Classification of the given points")
  if (labels) {legend("topright", legend = c(TeX("Points in $R_1$"), TeX("Points in $R_2$")), col = c('blue', 'red'), pch = 1)}
  return(collected_data)}
```

